---
title: "Earthquake Visualisation"
author: "Jamas Enright"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Earthquake Visualisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This demonstrates how to use the RCapstone package.

Start by loading in the data.

```{r, results='hide'}
library(dplyr)
raw_data<-readr::read_tsv(system.file("extdata", "signif.txt", package = "RCapstone"))
```

Then we can clean up the columns using eq_clean_data and eq_location_clean.

```{r}
clean_data<-eq_clean_data(raw_data)
clean_data<-eq_location_clean(clean_data)
```

We can now visualise this data two different ways.

One is as a line plot of earthquakes over time.

```{r}
subset(clean_data,COUNTRY %in% c("USA","CHINA")) %>%
  dplyr::filter(!is.na(EQ_PRIMARY), !is.na(DEATHS)) %>%
  ggplot2::ggplot() +
  ggplot2::aes(
    x = DATE,
    size = EQ_PRIMARY,
    colour = DEATHS,
    group=COUNTRY
  ) +
  geom_timeline()
```

And can add labels for the top sized earthquakes. Note that it will match any earthquakes to the top n values.

```{r}
subset(clean_data,COUNTRY %in% c("USA","CHINA")) %>%
  dplyr::filter(!is.na(EQ_PRIMARY), !is.na(DEATHS)) %>%
  ggplot2::ggplot() +
  ggplot2::aes(
    x = DATE,
    size = EQ_PRIMARY,
    colour = DEATHS,
    group=COUNTRY,
    label=LOCATION_NAME
  ) +
  geom_timeline() + 
  geom_timeline_label()
```

The other way is spatially. We can start with a basic map.

```{r}
clean_data %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
  eq_map(annot_col = "DATE")
```

We can improve this with better annotations.

```{r}
clean_data %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
  dplyr::mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = "popup_text")
```

